\Title

### 2021-04-28 latex/paper/main.tex
Finished intro, problem formulation, vnhc, and acrobot constraint sections. Need simulations, experiments, proof, and conclusion.

%/========== Main Document ==========/%
\begin{document}
\title{\Title}
\author{Adan Moran-MacDonald, \IEEEmembership{Member, IEEE}, Manfredi Maggiore,
\IEEEmembership{Member, IEEE}*, and Xingbo Wang
\thanks{Manuscript submitted for review on \today.}
\thanks{A. Moran-MacDonald is with the Department of Electrical and Computer
    Engineering, University of Toronto, Toronto, ON, Canada (e-mail:
adan.moran@mail.utoronto.ca).}
\thanks{M. Maggiore is with the Department of Electrical and Computer
Engineering, University of Toronto, ON, Canada (e-mail:
maggiore@control.utoronto.ca).}
\thanks{X. Wang is with ??? (e-mail: ???).}
} %/author

@@

%/========== Introduction ==========/%
\section{Introduction}\label{sec:introduction}
\IEEEPARstart{T}ODO: intro.

%/========== Problem Formulation ==========/%
\section{Problem Formulation}\label{sec:problem-formulation}
%TODO: motivate injection/dissipation for energy regulation in an acrobot. Use
%this to lead into VNHCs. e.g. represent torso as chain of a swing, pivot as
%seat, and legs as human legs. Replicate fig3.3 from thesis but for acrobot: take
%(qu,pu)-plane and show how a person moves their legs wrt theta, qu, or pu. Leave
%the gymnast model until the end, or ignore it entirely.
%
%Alternatively, ff human movement journal says legs move wrt qu, we can use that
%as a foundation.
%Ideas: Human movement says appropriate regulation of pendulum allows energy to
%be pumped into system. Pendulum length rapidly increased on downward swing, then
%rapidly decreased on upward swing (I think these mean backward and forward).
%They plot pendulum length rdot/r vs "body angle" theta to determine energy
%injection, which hints at some relationship between (r,rdot) and (theta,thetad).
%Indeed, the peak value of rdot/r did not occur at a fixed value of theta, nor
%once anyone reached a particular speed; it occured at the same thetadot/theta for all
%participants of the study. This perfectly motivates VNHCs for robots, because we
%can shape (r,rtdot) as functions of (theta,thetad).
%Start motivation from energy injection - how do humans inject energy? How does
%acrobot model gymnast? use human movement to motivate tracking functions of both
%position and velocity rather than time.

@@

\textbf{MANFREDI: This section feels more like an intro than a problem
formulation. Thoughts?}

@@

\subsection{Simply Actuated Hamiltonian Systems}
Take a mechanical system modelled with generalized coordinates 
\(q = (q_1, \ldots, q_n)\) on a configuration manifold
\(\mathcal{Q} = \Rt{T_1} \times \cdot \Rt{T_n}\), where
\(T_i = 2\pi\) if \(q_i\) is an angle and \(T_i = \infty\) if \(q_i\) is a
displacement. The corresponding generalized velocities are 
\(\dot{q} = (\dot{q}_1,\ldots,\dot{q}_n) \in \R^n\).

@@

It is easy to show using the matrix Kronecker product that
\eqref{eqn:hamiltonian-eom-general} expands to
\begin{equation}\label{eqn:hamiltonian-full-dynamics}
     \begin{cases}
        \dot{q} = D\inv(q)p 
        , \\
        \dot{p} = -\frac{1}{2} (\Id{n} \otimes p\tpose) \nabla_q D\inv(q) p
        - \nabla_q P(q) + B(q) \tau
        . \\
    \end{cases}
\end{equation}

@@

\subsection{Virtual Nonholonomic Constraints}

@@

We conclude this section by formalizing the notion of energy injection for
VNHCs.

Any system satisfying Definition \ref{defn:energy-gain}
can have unstable equilibria on \(D\), but not limit cycles nor closed orbits.
The next definition ties this notion of energy gain to VNHCs.

\begin{defn}\label{defn:energy-injection}
    A regular VNHC \(h(q, p) = 0\) with constraint manifold \(\Gamma\)
    \textit{injects (dissipates) energy on \(D \subset \Gamma\)} if the
    constrained dynamics gain (lose) energy everywhere on \(D\), except possibly
    on a set of measure zero.
\end{defn}

@@

%/========== Acrobot ==========/%
\section{The Acrobot VNHC}\label{sec:acrobot}
%specialize the theory of VNHCs to the acrobot. consider constraints which depend
%only on pu, and show arctan vnhc and reduced dynamics. 
%give definition of energy injection/dissipation. conclude with our
%theorem. Make this section short and sweet.
%
%TODO: after writing S2, re-introduce what we want to do with acrobot (energy
%injection for giant stabilization)
Our goal in this article is to design a VNHC which injects energy into the
acrobot through giant-like motion.
We will use the simplified acrobot model in Figure
\ref{fig:simple-acrobot-model}, where we assume the torso and leg rods are of
equal length \(l\) with equal point masses \(m\) at the tips.
A real gymnast cannot swing their legs in full circles, though they
are usually flexible enough to raise them parallel to the floor;
hence, we assume the leg angle \(q_a\) lies in \([-Q_a, Q_a]\) for some
\(Q_a \in [\frac{\pi}{2}, \pi[\). 
We also ignore any frictional forces.

\begin{figure}
    \centering
    \includegraphics[width=0.7\linewidth]{simple_acrobot_model.pdf}
    \caption{A simple acrobot has massless rods of equal length \(l\) and 
    equal masses \(m\) at the tips.}
    \label{fig:simple-acrobot-model}
\end{figure}

The acrobot has inertia matrix 
\(M\), potential function \(V\) (with respect to
the horizontal bar), and input matrix \(B\) given as follows:
\begin{align}\label{eqn:acrobot-inertia}
    M(q) &= \begin{bmatrix}
        ml^2\left(3+2\cos(q_a)\right) & 
        ml^2\left(1+\cos(q_a)\right) \\
        ml^2\left(1+\cos(q_a)\right) &
        ml^2
    \end{bmatrix} 
    , \\
    \label{eqn:acrobot-potential}
    V(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
    , \\
    \label{eqn:acrobot-B}
    B &= [0;1]
    .
\end{align}
The conjugate of momenta is \(p = (p_u,p_a) = M(q)\dot{q}\).
The dynamics of the acrobot in \((q,p)\) coordinates are given in
~\eqref{eqn:acrobot-hamiltonian}, where
for shorthand, we write \(c_u := \cos(q_u)\), \(c_a := \cos(q_a)\), and 
\(c_{ua} := \cos(q_u + q_a)\); likewise, \(s_u := \sin(q_u)\), 
\(s_a := \sin(q_a)\), and \(s_{ua} := \sin(q_u + q_a)\).
\begin{align}\label{eqn:acrobot-hamiltonian}
    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
    mgl\left(2 c_u + c_{ua}\right)
    , \\
     &\begin{cases}
        \dot{q} = \Minv(q) p 
        ,\\
        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) 
        ,\\
        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
        - mgl s_{ua} + \tau,
    \end{cases} \nonumber
\end{align}
The control input is a force \(\tau \in \R\) affecting only the dynamics of
\(p_a\), representing a torque acting on the hip joint.
This means \((q,p)\) are simply actuated coordinates in the phase space
\(\mathcal{Q} \times \mathcal{P}\) where
\(\mathcal{Q} = \Sone \times \Sone\), and
\(\mathcal{P} = \R \times \R\).
We can therefore apply the theory from Section \ref{sec:vnhc} to design
a VNHC of the form \(q_a = f(q_u,p_u)\). 
Since we need the VNHC to be regular, the following proposition will be
useful.
\begin{prop}\label{prop:acrobot-fpu-regular}
    A relation \(h(q,p) = q_a - f(p_u) = 0\) 
    with \(f \in C^2\left(\R; \Sone\right)\) is a regular
    VNHC of order 1 for the acrobot.
\end{prop}
\begin{proof}
    The regularity matrix for the acrobot evaluates to
    \(((1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a))/(ml^2(2-c_a^2))\).
    Since \(\partial_{q_u} f = 0\), this matrix is strictly positive for all values
    of \(q_a\), and hence is full rank everywhere on the constraint manifold.
\end{proof}

@@

Recall that \((q_u, p_u)\) denote the angle and momentum of the acrobot's torso.
By Theorem \ref{thm:zero-dynamics}, the constrained acrobot is parameterized
fully by \((q_u,p_u) \in \SxR\).
It is easy to show that the constrained dynamics are
just the torso dynamics obtained when one swings the legs
according to \eqref{eqn:acrobot-constraint}:
\begin{equation}\label{eqn:acrobot-constrained-dynamics}
    \begin{cases}
    \dot{q}_u &= \frac{(1+I^2 p_u^2)p_u + m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }
            {ml^2(1+I^2 p_u^2)(3+2c_a)}
    ,    \\
    \dot{p}_u &= - m g l (2s_u + s_{ua})
    ,
    \end{cases}
\end{equation}
subject to \(q_a = \bar{q}_a \arctan(I p_u)\).

@@

%/========== Experiments ==========/%
\section{Experimental Results}\label{sec:experiments}
We tested our VNHC on the physical acrobot built by
Xingbo Wang \cite{xingbo_thesis} (Figure \ref{fig:xingbo-acrobot}).
The torso and legs on this robot have differing masses and lengths, so
its inertia matrix and potential function are \textbf{not} described by 
~\eqref{eqn:acrobot-inertia}-\eqref{eqn:acrobot-potential}.
If our VNHC injects energy into this acrobot, then we can infer that
VNHCs are robust to model mismatch.

\begin{figure}
    \centering
    \includegraphics[width=0.6\linewidth]{xingbo_acrobot.jpg}
    \caption{The acrobot built by Wang \cite{xingbo_thesis}.}
    \label{fig:xingbo-acrobot}
\end{figure}

@@

\begin{figure}[]
    \centering
    \includegraphics[width=0.7\linewidth]{acrobot_orbit.png}
    \caption{A simulated orbit of Wang's acrobot.}
    \label{fig:acrobot-orbit}
\end{figure}
To verify numerically that the acrobot would consistently achieve rotations, we
ran a Monte-Carlo simulation \cite{montecarlo} where we initialized the acrobot
randomly inside the sublevel set
\[
    \left\{(q_u,p_u) \in \SxR \mid
    E(q_u,p_u) \leq E\left(\frac{\pi}{32},0\right)\right\}
    ,
\] 
and measured the how long it took to begin rotating.
The results in Figure \ref{fig:acrobot-mc} show that
the acrobot always rotated within 10--35 seconds.

\begin{figure}[]
    \centering
    \includegraphics[width=0.7\linewidth]{acrobot_mc.png}
    \caption{Monte Carlo simulation results.}
    \label{fig:acrobot-mc}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.45\linewidth}
        \includegraphics[width=\linewidth]{acrobot_unperturbed_orbit.png}
        \caption{The baseline test.}
        \label{fig:acrobot-unperturbed-orbit}
    \end{subfigure}
    \begin{subfigure}[t]{0.45\linewidth}
        \includegraphics[width=\linewidth]{acrobot_stopped_orbit.png}
        \caption{Acrobot orbit before (blue) and after (red) stopping.}
        \label{fig:acrobot-stopped-orbit}
    \end{subfigure}
    \caption{The physical acrobot's orbit during baseline and stop tests.}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[ht]{0.45\linewidth}
        \includegraphics[width=\linewidth]{acrobot_fpush_orbit.png}
        \caption{The forwards push test.}
        \label{fig:acrobot-fpush-orbit}
    \end{subfigure}
    \begin{subfigure}[ht]{0.45\linewidth}
        \includegraphics[width=\linewidth]{acrobot_rpush_orbit.png}
        \caption{The reverse push test.}
        \label{fig:acrobot-rpush-orbit}
    \end{subfigure}
    \caption{The acrobot orbit before (blue) and after (red) pushing.}
\end{figure}

\subsection{Summary of Results} 
These simulations and experiments demonstrate that VNHCs are excellent
tools for injecting energy.
In particular, energy injection through VNHCs appears to be robust against a
variety of disturbances, including significant model mismatch.

%/========== Proof ==========/%
\section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:proof}

>>>>

%/========== Main Document ==========/%
\begin{document}
\title{\Title}
\author{Adan Moran-MacDonald, \IEEEmembership{Member, IEEE}, Manfredi Maggiore,
\IEEEmembership{Member, IEEE}*, and Xingbo Wang
\thanks{Manuscript submitted for review on \today.}
\thanks{A. Moran-MacDonald (e-mail: adan.moran@mail.utoronto.ca) and
M. Maggiore (e-mail: maggiore@control.utoronto.ca) are with the Department of
Electrical and Computer Engineering, University of Toronto, ON, Canada.}
\thanks{X. Wang is with ??? (e-mail: ???).}
} %/author

@@

%/========== Introduction ==========/%
\section{Introduction}\label{sec:introduction}

@@

%/========== Problem Formulation ==========/%
\section{Problem Formulation}\label{sec:problem-formulation}
We will use the simplified acrobot model in Figure
\ref{fig:simple-acrobot-model}, where we assume the torso and leg rods are of
equal length \(l\) with equal point masses \(m\) at the tips.
A real gymnast cannot swing their legs in full circles, though they
are usually flexible enough to raise them parallel to the floor;
hence, we assume the leg angle \(q_a\) lies in \([-Q_a, Q_a]\) for some
\(Q_a \in [\frac{\pi}{2}, \pi[\). 
We also ignore any frictional forces.

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{simple_acrobot_model.pdf}
    \caption{A simple acrobot has massless rods of equal length \(l\) and 
    equal masses \(m\) at the tips.}
    \label{fig:simple-acrobot-model}
\end{figure}

The acrobot has inertia matrix 
\(M\), potential function \(V\) (with respect to
the horizontal bar), and input matrix \(B\) given as follows:
\begin{align}\label{eqn:acrobot-inertia}
    M(q) &= \begin{bmatrix}
        ml^2\left(3+2\cos(q_a)\right) & 
        ml^2\left(1+\cos(q_a)\right) \\
        ml^2\left(1+\cos(q_a)\right) &
        ml^2
    \end{bmatrix} 
    , \\
    \label{eqn:acrobot-potential}
    V(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
    , \\
    \label{eqn:acrobot-B}
    B &= [0;1]
    .
\end{align}
For reasons that will become clear in later sections of this article, we will
use Hamiltonian mechanics to derive the dynamics of the acrobot. 
For this we require the conjugate of momenta, \(p = (p_u,p_a) = M(q)\dot{q}\).
The dynamics of the acrobot in \((q,p)\) coordinates are given in
~\eqref{eqn:acrobot-hamiltonian}.
For shorthand, we write \(c_u := \cos(q_u)\), \(c_a := \cos(q_a)\), and 
\(c_{ua} := \cos(q_u + q_a)\); likewise, \(s_u := \sin(q_u)\), 
\(s_a := \sin(q_a)\), and \(s_{ua} := \sin(q_u + q_a)\).
\begin{align}\label{eqn:acrobot-hamiltonian}
    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
    mgl\left(2 c_u + c_{ua}\right)
    , \\
     &\begin{cases}
        \dot{q} = \Minv(q) p 
        ,\\
        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) 
        ,\\
        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
        - mgl s_{ua} + \tau.
    \end{cases} \nonumber
\end{align}
The control input is a force \(\tau \in \R\) affecting only the dynamics of
\(p_a\), representing a torque acting on the hip joint.

Our goal is to design a smooth function \(f : \Sone \times \R\) such that the
relation \(q_a = f(q_u,p_u)\) is a regular VNHC, where the constrained dynamics
of the acrobot (under this constraint) gain energy on some set \(D\)
in the sense of Definition \ref{defn:energy-gain}.
We will formally define regularity in the next section.
Essentially, if the constraint meets some regularity conditions then it can be
stabilized via feedback.

Note that any system satisfying Definition \ref{defn:energy-gain}
can have unstable equilibria on \(D\), but not limit cycles nor closed orbits.

@@

\subsection{Simply Actuated Hamiltonian Systems} \label{sec:vnhc-sah}
Take a mechanical system modelled with generalized coordinates 
\(q = (q_1, \ldots, q_n)\) on a configuration manifold
\(\mathcal{Q} = \Rt{T_1} \times \cdot \Rt{T_n}\), where
\(T_i = 2\pi\) if \(q_i\) is an angle and \(T_i = \infty\) if \(q_i\) is a
displacement. The corresponding generalized velocities are 
\(\dot{q} = (\dot{q}_1,\ldots,\dot{q}_n) \in \R^n\).

@@

It is easy to show using the matrix Kronecker product that
\eqref{eqn:hamiltonian-eom-general} expands to
\begin{equation*}\label{eqn:hamiltonian-full-dynamics}
     \begin{cases}
        \dot{q} = D\inv(q)p 
        , \\
        \dot{p} = -\frac{1}{2} (\Id{n} \otimes p\tpose) \nabla_q D\inv(q) p
        - \nabla_q P(q) + B(q) \tau
        . \\
    \end{cases}
\end{equation*}
%\begin{align*}
%    \dot{q} &= D\inv(q)p 
%    ,\\
%    \dot{p} &= -\frac{1}{2} (\Id{n} \otimes p\tpose) \nabla_q D\inv(q) p
%        - \nabla_q P(q) + B(q) \tau
%    .
%\end{align*}

@@

\subsection{Virtual Nonholonomic Constraints}\label{sec:vnhc-vnhc}

@@

We conclude this section by formalizing the notion of energy
injection/dissipation for VNHCs.

\begin{defn}\label{defn:energy-injection}
    A regular VNHC \(h(q, p) = 0\) with constraint manifold \(\Gamma\)
    \textit{injects (dissipates) energy on \(D \subset \Gamma\)} if the
    constrained dynamics gain (lose) energy everywhere on \(D\) according to
    Definition \ref{defn:energy-gain}, except possibly on a set of measure
    zero.
\end{defn}

@@

%/========== Acrobot ==========/%
\section{The Acrobot VNHC}\label{sec:acrobot}
%specialize the theory of VNHCs to the acrobot. consider constraints which depend
%only on pu, and show arctan vnhc and reduced dynamics. 
%give definition of energy injection/dissipation. conclude with our
%theorem. Make this section short and sweet.
%
%TODO: after writing S2, re-introduce what we want to do with acrobot (energy
%injection for giant stabilization)
Our goal in this article is to design a VNHC which injects energy into the
acrobot through giant-like motion.
Recall that the acrobot in Figure \ref{fig:simple-acrobot-model} has dynamics given by
~\eqref{eqn:acrobot-hamiltonian}, which are
repeated here:
\begin{align*}
    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
    mgl\left(2 c_u + c_{ua}\right)
    , \\
     &\begin{cases}
        \dot{q} = \Minv(q) p 
        ,\\
        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) 
        ,\\
        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
        - mgl s_{ua} + \tau.
    \end{cases}
\end{align*}

Since the control input \(\tau\) only affects the actuated momentum,
\((q,p)\) are simply actuated coordinates in the phase space
\(\mathcal{Q} \times \mathcal{P}\) where
\(\mathcal{Q} = \Sone \times \Sone\), and
\(\mathcal{P} = \R \times \R\).
We can therefore apply the theory from Section \ref{sec:vnhc} to design
a VNHC of the form \(q_a = f(q_u,p_u)\). 
Since we need the VNHC to be regular, the following proposition will be
useful.
\begin{prop}\label{prop:acrobot-fpu-regular}
    A relation \(h(q,p) = q_a - f(p_u) = 0\) 
    with \(f \in C^2\left(\R; \Sone\right)\) is a regular
    VNHC of order 1 for the acrobot.
\end{prop}
\begin{proof}
    The regularity matrix for the acrobot evaluates to
    \(((1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a))/(ml^2(2-c_a^2))\).
    Since \(\partial_{q_u} f = 0\), this matrix is strictly positive for all values
    of \(q_a\), and hence is full rank everywhere on the constraint manifold.
\end{proof}

@@

Recall that \((q_u, p_u)\) denote the angle and momentum of the acrobot's torso.
By Theorem \ref{thm:zero-dynamics}, the constrained acrobot is parameterized
fully by \((q_u,p_u) \in \SxR\).
It is easy to show that the constrained dynamics are just the torso dynamics
obtained when one swings the legs according to \eqref{eqn:acrobot-constraint}.
That is, the constrained dynamics are
\begin{equation}\label{eqn:acrobot-constrained-dynamics}
    \begin{cases}
    \dot{q}_u &= \frac{(1+I^2 p_u^2)p_u + m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }
            {ml^2(1+I^2 p_u^2)(3+2c_a)}
    ,    \\
    \dot{p}_u &= - m g l (2s_u + s_{ua})
    ,
    \end{cases}
\end{equation}
subject to \(q_a = \bar{q}_a \arctan(I p_u)\).

@@

The results of Theorem \ref{thm:acrobot-energy-stabilization} lead naturally to
the idea of energy regulation: one could employ a supervisor which switches
between injection and dissipation constraints to stabilize a desired energy
level \(\bar{E}\).
For instance, the supervisor could inject energy when \(E(q_u,p_u) < \bar{E}\) 
and dissipate energy when \(E(q_u,p_u) > \bar{E}\), with some hysteresis to
avoid infinite switching.
As we shall see, such a simple supervisor will not produce human-like
oscillations or rotations.
However, it will allow the acrobot to stabilize some peak angle of oscillation
\(\bar{q}_u \in \, ]0,\pi[\) so that \(E(\pm \bar{q}_u,0) = \bar{E}\),
or some peak rotational momentum
\(\bar{p}_u \in \, ]\sqrt{60m^2gl^3}, \bar{\rho}[\) where
\(E(0,\pm \bar{p}_u) = \bar{E}\).

%/========== Experiments ==========/%
\section{Simulation Results}\label{sec:simulations}
\textbf{TODO: simulation parameters in a table, show results for injection,
dissipation, and regulation along with existing simulations. Ample commentary.}

\section{Experimental Results}\label{sec:experiments}
\textbf{TODO: 2 sections. a) is hardware description, paraphrasing Xingbo.
    Should be brief but detailed (eg 2 arduinos, 3d printed, battery in the leg,
    parameters of the system, etc). b) is
experimental results, same as simulation section.}
We tested our VNHC on the physical acrobot built by
Xingbo Wang \cite{xingbo_thesis} (Figure \ref{fig:xingbo-acrobot}).
The torso and legs on this robot have differing masses and lengths, so
its inertia matrix and potential function are \textbf{not} described by 
~\eqref{eqn:acrobot-inertia}-\eqref{eqn:acrobot-potential}.
If our VNHC injects energy into this acrobot, then we can infer that
VNHCs are robust to model mismatch.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{xingbo_acrobot.jpg}
    \caption{The acrobot built by Wang \cite{xingbo_thesis}.}
    \label{fig:xingbo-acrobot}
\end{figure}

@@

\begin{figure}[]
    \centering
    \includegraphics[width=0.8\linewidth]{acrobot_orbit.png}
    \caption{A simulated orbit of Wang's acrobot.}
    \label{fig:acrobot-orbit}
\end{figure}
To verify numerically that the acrobot would consistently achieve rotations, we
ran a Monte-Carlo simulation \cite{montecarlo} where we initialized the acrobot
randomly inside the sublevel set
\[
    \left\{(q_u,p_u) \in \SxR \mid
    E(q_u,p_u) \leq E\left(\frac{\pi}{32},0\right)\right\}
    ,
\] 
and measured the how long it took to begin rotating.
The results in Figure \ref{fig:acrobot-mc} show that
the acrobot always rotated within 10--35 seconds.

\begin{figure}[]
    \centering
    \includegraphics[width=0.8\linewidth]{acrobot_mc.png}
    \caption{Monte Carlo simulation results.}
    \label{fig:acrobot-mc}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_unperturbed_orbit.png}
        \caption{The baseline test.}
        \label{fig:acrobot-unperturbed-orbit}
    \end{subfigure}
    \begin{subfigure}[t]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_stopped_orbit.png}
        \caption{Acrobot orbit before (blue) and after (red) stopping.}
        \label{fig:acrobot-stopped-orbit}
    \end{subfigure}
    \caption{The physical acrobot's orbit during baseline and stop tests.}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[ht]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_fpush_orbit.png}
        \caption{The forwards push test.}
        \label{fig:acrobot-fpush-orbit}
    \end{subfigure}
    \begin{subfigure}[ht]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_rpush_orbit.png}
        \caption{The reverse push test.}
        \label{fig:acrobot-rpush-orbit}
    \end{subfigure}
    \caption{The acrobot orbit before (blue) and after (red) pushing.}
\end{figure}

\subsection{Summary of Results} 
These simulations and experiments demonstrate that VNHCs are excellent
tools for injecting energy.
In particular, energy injection through VNHCs appears to be robust against a
variety of disturbances, including significant model mismatch and sensor noise.

%/========== Proof ==========/%
\section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:proof}
\textbf{TODO: prove Thm 10}
